---
title: "Reproducible Research, Mid-course Project"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(ggplot2)
dat <- read.csv("activity.csv",header=TRUE)
```

##What is the mean total number of steps taken per day?
```{r}
s <- with(dat,aggregate(steps~date,FUN=sum))
qplot(s$steps,geom="histogram")
```

This is a histogram of the total number of steps taken per day. The mean number of steps taken is `r as.integer(mean(s$steps))`, and the median number of steps taken is `r median(s$steps)`.

##What is the average daily activity pattern?
```{r}
m <- with(dat,aggregate(steps~interval,FUN=mean))
qplot(m$interval,m$steps)
```

This is a time series of the average number of steps taken during each time interval. The interval with the maximum steps taken on average is interval `r m$interval[which(m$steps==max(m$steps))]`.

##Imputing missing values

```{r}
nas <- sum(is.na(dat$steps))
```

There are `r nas` missing values. To fill in the missing values, we can replace them with the average number of steps taken in each interval.

```{r}
nas_filled <- dat
for(i in 1:length(nas_filled$steps)){
  if(is.na(nas_filled$steps[i])){
    interval <- nas_filled$interval[i]
    avg_steps <- m$steps[which(m$interval==interval)]
    nas_filled$steps[i] <- avg_steps
  }
}
new_s <- with(nas_filled,aggregate(steps~date,FUN=sum))
qplot(new_s$steps,geom="histogram")
```

The new mean steps per day is `r as.integer(round(mean(new_s$steps)))`, and the new median steps per day is `r as.integer(round(median(new_s$steps)))`. This is not significantly different than before the adjustment. By adding these missing values, the curve of the histogram plot becomes more smoothed-out. 

##Are there differences in activity patterns between weekdays and weekends?

```{r}
nas_filled$weekday <- wday(nas_filled$date)
for(i in 1:length(nas_filled$weekday)){
  if(nas_filled$weekday[i]==1){
    nas_filled$weekday[i] <- "Weekend"
  }
  else if(nas_filled$weekday[i]==7){
    nas_filled$weekday[i] <- "Weekend"
  }
  else{
    nas_filled$weekday[i] <- "Weekday"
  }
}
#split data into weekdays,weekends
#aggregate by interval
dat_weekdays <- nas_filled[which(nas_filled$weekday=="Weekday"),]
dat_weekend <- nas_filled[which(nas_filled$weekday=="Weekend"),]
avg_weekdays <- with(dat_weekdays,aggregate(steps~interval,FUN=mean))
avg_weekend <- with(dat_weekend,aggregate(steps~interval,FUN=mean))
par(mfrow=c(1,2))
plot(avg_weekdays$interval,avg_weekdays$steps)
plot(avg_weekend$interval,avg_weekend$steps)

```

There are some differences in activity during weekdays and weekends.



